---
title: "Practical Machine Learning Course Project"
author: "JLara"
date: "23 de julio de 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: <http://groupware.les.inf.puc-rio.br/har> (see the section on the Weight Lifting Exercise Dataset).

## Data

The training data for this project are available here:

<https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv>

The test data are available here:

<https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv>


The data for this project come from this source: <http://groupware.les.inf.puc-rio.br/har>. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.

# Loading and preprocessing the data

* Load the data

```{r load}

library(caret)
library(randomForest)
library(gbm)
library(AppliedPredictiveModeling)
library(rpart)
library(plyr)

set.seed(666)

trainURL <- 'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv'

data_train <- "./pml-training.csv"

testURL <- 'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv'

data_test <- "./pml-testing.csv"

# Download the file
if (file.exists(trainURL) == FALSE) {
  download.file(trainURL,destfile=data_train)
}

if (file.exists(testURL) == FALSE) {
  download.file(testURL,destfile=data_test)
}
# Read the training data into R, identifying “NA”, “” and “#DIV/0!” as NA strings

train_d <- read.csv("pml-training.csv",na.strings=c("NA","","#DIV/0!"))
test_d <- read.csv("pml-testing.csv",na.strings=c("NA","","#DIV/0!"))

```


```{r str,results='hide'} 
str(train_d)
str(test_d)
```

We have for the data frame train 19622 observations of 160 variables and for the data frame test we have 20 observations of 160 variables
Given the number of NA we have, we eliminate the variables that have everything with NA.

```{r sinNA} 
sinNAtra <- sapply(train_d, function(x)all(!is.na(x)))
sinNAtes <- sapply(test_d, function(x)all(!is.na(x)))
train_d <- train_d[,sinNAtra]
test_d <- test_d[,sinNAtes]

str(train_d)
str(test_d)

# Delete variables are irrelevant to our current project: user_name, raw_timestamp_part_1, raw_timestamp_part_,2 cvtd_timestamp, new_window, and  num_window (columns 1 to 7). 
train_d <-train_d[,-c(1:7)]
test_d  <-test_d[,-c(1:7)]

```

```{r strnew,results='hide'} 
str(train_d)
str(test_d)
```

We have for the data frame train 19622 observations of 160 variables and pra the data frame test we have 20 observations of 160 variables Given the number of NA we have, we eliminate the variables that have everything with NA.



## Model Building

For the project we will use three models (decision trees with CART, stochastic gradient boosting tress and random forest decision trees) to see quient has the best accuarity. On the best we show details cross validation.


Partition the train data so that 70% of the training dataset into training and the remaining 30% to testing.

```{r partition}
trainset <- createDataPartition(y=train_d$classe, p=0.7, list=FALSE)
training <- train_d[trainset, ] 
testing <- train_d[-trainset, ]
```

We evaluate the models.

```{r model}

controls <- trainControl(method = "cv", number = 6)
mod_rf <- train(classe ~ ., data = training, method = "rf",trControl=controls)
mod_gbm <- train(classe ~ ., data = training, method = "gbm",trControl=controls)
mod_rpart <- train(classe ~ ., data = training, method = "rpart",trControl=controls)
pred_rf <- predict(mod_rf, testing)
pred_gbm <- predict(mod_gbm, testing)
pred_rpart <- predict(mod_rpart, testing)
confusionMatrix(pred_rf, testing$classe)$overall[1]
confusionMatrix(pred_gbm, testing$classe)$overall[1]
confusionMatrix(pred_rpart, testing$classe)$overall[1]

```

The best model is Random Forest which gives us a higher Accuracy with 0.9924. We will use the Random Forest model to make the prediction.

```{r cross}
confusionMatrix(pred_rf, testing$classe)
```

Accuracy for Random Forest model was 0.9924 (95% CI: (0.9898, 0.9944)). The expected out-of-sample error is estimated at 0.0076, or 0.76%.

## Prediction using the model

Here is the final outcome based Random Forest applied against the Testing dataset.

```{r prediction}
prediction <- predict(mod_rf,test_d)
print(prediction)
```